Towards typed repositories of proofs

### 2010-06-18 papers/MIPS2010/abstract.tex
passe sur le kernel
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@120 403e655e-980a-47bd-8c60-1ee5348bb6b7

Representing syntax and logics is nicely done in a \emph{logical
  framework} like LF: both the syntactic
elements and the typing derivations can sit in the same tree
structure, and both can be rechecked at the same time, thanks to
dependent types. For the purpose of incremental type checking though,
our needs are a little bit different: first, we need to record, that
is to name all intermediate values of our developments, so as to be
able to address and reuse them multiple times. Secondly, we need to
make sure that those intermediate values (sub-terms) are not recorded
twice, so as to not type-check them twice: we are looking to represent
syntactic and typing objects as a DAG rather than a tree.

Our system will have these properties w.r.t. LF:
\begin{itemize}
\item In this first iteration of the project, we do not need
  computations to take place within our DAGs. Our syntax will then be
  restricted to product types and applications;
\item Every term should be a flat application of variables, so that we
  don't introduce compound terms without naming them;
\item Finally we need a way to record intermediate definitions: we
  introduce a new kind of binder, the equality binder. We maintain the
  invariant that not two same definition can sit in the context while
  typing.
\end{itemize}

We describe the state of a repository at a given moment by a type in
the following syntax. Well-typed types in the repository meta-language
guarantee that it contains only well-typed proofs in the object
language.
\begin{align*}
 \gdecl{a}{x \gor a\ x } \\
 \gdecl{t}{a \gor \sort \gor (x:t)\cdot t \gor (x=a:t)\cdot t} \\
 \gdecl{\Gamma}{\cdot \gor \Gamma[x:t] \gor \Gamma[x=a:t]}
\end{align*}

\paragraph{Related work}
\label{rw}

>>>>

Representing syntax and logics is nicely done in a \emph{logical
  framework} like LF: both the syntactic elements and the typing
derivations can sit in the same tree structure, and both can be
rechecked at the same time, thanks to dependent types. For the purpose
of incremental type checking though, our needs are a little bit
different: first, we need to record, that is to name all intermediate
values of our developments, so as to be able to address and reuse them
multiple times. Secondly, we need to make sure that those intermediate
values (sub-terms) are not recorded twice, so as to not type-check
them twice: we are looking to represent syntactic and typing objects
as a DAG rather than a tree. Moreover, we enforce by typing a property
of \emph{maximal sharing}: every different subterm can be constructed
exactly once. Our system will have these properties w.r.t. LF:
\begin{itemize}
\item In this first iteration of the project, we do not need
  computations to take place within our DAGs. Our syntax will then be
  restricted to product types and applications;
\item Every term should be a flat application of variables, so that we
  don't introduce compound terms without naming them and recording
  their types;
\item Finally we need a way to record intermediate definitions: we
  introduce a new kind of binder, the equality binder. We maintain the
  invariant that not two same definition can sit in the context while
  typing, thus guaranteeing \emph{maximal sharing} among terms and
  derivations.
\end{itemize}

We describe the state of a repository at a given moment by a type $t$
in the following syntax. Well-typed types in the repository
meta-language guarantee that it contains only well-typed proofs in the
object language.
\begin{align*}
 \gdecl{a}{x \gor a\ x } \\
 \gdecl{t}{a \gor \sort \gor (x:t)\cdot t \gor (x=a:t)\cdot t}
\end{align*}

From an implementation point of view, following \textsf{GIT}, we store
these terms in a database of objects indexed by \emph{keys}, which are
hash values of their contents, so that retreiving the type of a whole
term boils down to compute its hash, and look for it in the database.

Note that in this system, computation has no existence: we only
provide ways to verify that syntax and typing rules of an object
language are correct. Therefore, patches -- functions from
repositories to repositories -- have no existence \emph{per se}. A
necessary enhancement of our theory would be to consider computation
as a way to apply patches as if they were constructive
metatheorems. We conjecture that it would be done by re-extending the
system towards LF (adding abstraction and reduction).

\paragraph{Related works}
\label{rw}
